### 2012-09-13 relwork-scratch.tex
relwork and bibtex

extraction/collection from pages
                      sifter
                      hunter-gatherer
                      cards
     TX2 - learning during collection

post-scraping structured data mashups
    pipes, popfly, QDEWiki

PIM platform (on the fly data integration)
    SEMEX - 
     Haystack -

mediating schema, pairwise mediationnz

scraperwiki

----

As the theoretical basis of data integration extends nearly a half-century in the database systems literature . As described by Halvey et al’s comprehensive textbook on data integration problems and techniques \cite{principlesofdataintegration} much of this focus of this community has focused on \emph{engineered} or \emph{automatic} approaches to data integration at a large scale,

In scientific laboratory settings,

We only briefly touch on the extensive literature covering the “deep end” of data integration - both automatic and “hand-coded” techniques to link multiple, typically large, data sources in consistent, correct and efficient ways, little work has pertained on the opposite end -- the `light', “small” and “sketchy” end of data integration, such as arises when multiple heterogeneous sources of data about similar kinds of things are brought together spontaneously

Interfaces that let end-users find, consolidate, and work with heterogeneous data 
PIM 
	Amy Voida’s databases paper
User-subjective approach to information management

Information scraps
End-user Data Gathering and integration platforms
	Potluck
	Cards (M. Dontcheva)

End-user data gathering
	Piggy Bank
“data notebooks”	Hunter-gatherer
	Yahoo Pipes
PROMPT

A key assumption of this work is that ontologies are not going to converge.  (Studies about the diversity of web ontologies backing this claim up?)

Why end-user data integration is desirable
User-subjective approach to personal information management

Why is the data integration problem so hard?  
The problem of reconciling schema heterogeneity has been a subject of research for decades, but solutions are few.  In the words of leading data integration researcher Alon Halevy,
[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
overlapping domains, they will model them in different ways. Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal. \cite{alonhalevy}

Exhibit - designed for personal structured curation
Yahoo Pipes! - designed for not-personal (public) curation
MySupermarket - http://mysupermarket.co.uk/
Demonstrates the power of joining across data sources: people can gather a grocery list, and compare how much the same shopping list would cost at each of the major UK supermarket chains (ASDA, Tesco, Waitrose, ocado) --

But what about the mom + pop groceries?
What if they’re missing data: allergies, suitable for -- available from somewhere else

Schema.org  - is it helping or not?